Experiment 4 of 4

### 2016-07-28 experiments/e4/writeup/e4.Rmd
e4 edits

```{r, startClean, echo=FALSE}
# start clean
rm(list=ls())
```

```{r, docOpts, echo=FALSE, cache=FALSE}
# Knitr doc options.
library(knitr)
knit_hooks$set(crop=hook_pdfcrop)
#options(tidy=TRUE)
#options(replace.assign=TRUE, width=50)
```

```{r, loadLibraries, echo=FALSE}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

@@

```{r, sourceFunctions}
source("functions/CIfunctions.R")
```

Get data from the raw per-subject files.

```{r, getData}
#if( file.exists('dd.Rda') ) load('dd.Rda') else {
#system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "../data/data.txt", sep=""))
#system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "../data/data.txt", sep=""))
dat <- read.delim(file="../data/data.txt", header=T, stringsAsFactors=FALSE)

@@

save(dat, file="dat.Rda")

dd <- dat

@@

# add transformations of previous RT
dd$RTprev_RecSqd    <- 1/dd$RTprev^2
dd$RTprev_Rec       <- 1/dd$RTprev
dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
dd$RTprev_log       <- log(dd$RTprev)
dd$RTprev_sqt       <- sqrt(dd$RTprev)
dd$RTprev_raw       <- dd$RTprev
dd$RTprev_sqd       <- dd$RTprev^2
  # save dd
  save(dd, file="dd.Rda")
#}
```

@@

```{r,fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("ggplot2"))
load("dat.Rda")

@@

temp <- melt(data=dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
             variable.name="transformation",
             value.name="score")

ggplot(temp, aes(score, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(dfSubset~transformation, scales="free", nrow=2) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```

@@

```{r, fig.width=20, fig.height=5}
# boxplot subjects and items

load("dd.Rda")
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

@@

```{r, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.

suppressMessages(library(plyr))
suppressMessages(library(lattice))
suppressMessages(library(latticeExtra))

subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

@@

```{r, message=F, results='hide'}
if(require(ggplot2)){}
load("dd.Rda")
```

```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

@@

```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

@@

```{r}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

@@

```{r, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

@@

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.01) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

@@

```{r}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Selection",4)
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Selection:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Selection:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
```

```{r,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

@@

```{r, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Selection")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Selection"
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity", "Selection")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Selection"
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Order","Selection")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Selection"
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"

@@

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + geom_line() + theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

@@

```{r, rtmodel-allitems}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lmer(
  data=dd,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm)
```

@@

```{r, rtmodeltable, results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

@@

```{r}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

@@

```{r}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber)
```

```{r, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

@@

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

@@

```{r}
dd2 <- subset(dd, abs(rstand)<3.5 )
```

```{r}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lmer(
  data=dd2,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm)
```

@@

```{r,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

@@

```{r}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

@@

```{r}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber
)

```{r, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

@@

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

@@

```{r,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='latex')
```

>>>>

```{r, docOpts, echo=FALSE, cache=FALSE}
library(knitr)
knit_hooks$set(crop=hook_pdfcrop)
```

```{r, loadLibraries, echo=FALSE}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(latticeExtra))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

@@

```{r, sourceFunctions}
source("CIfunctions.R")
```

Get data from the raw per-subject files if necessary.

```{r, getData}
if (file.exists('data.txt')) {
	dat <- read.delim(file="data.txt", header=T, stringsAsFactors=FALSE)
} else {
	system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "data.txt", sep=""))
	system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "data.txt", sep=""))
}

@@

@@

# add transformations of previous RT
dd$RTprev_RecSqd    <- 1/dd$RTprev^2
dd$RTprev_Rec       <- 1/dd$RTprev
dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
dd$RTprev_log       <- log(dd$RTprev)
dd$RTprev_sqt       <- sqrt(dd$RTprev)
dd$RTprev_raw       <- dd$RTprev
dd$RTprev_sqd       <- dd$RTprev^2

save(dd, file="data.Rda")
```

@@

```{r foo,fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

@@

temp <- reshape2::melt(data=dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"))

ggplot(temp, 
			 aes(value, colour=RTtype)) + 
	geom_freqpoly() + 
  facet_wrap(dfSubset~variable, scales="free", nrow=2) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```

@@

```{r bar, fig.width=20, fig.height=5}
# boxplot subjects and items
load("data.Rda")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw")) 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

@@

```{r, bar2, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

@@

```{r barr3, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

@@

```{r, var3, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

@@

```{r, bar4}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

@@

```{r, bar5, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

@@

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.01) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
```

@@

```{r, bar6}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Selection",4)
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Selection:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Selection:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
```

```{r,bar7,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

@@

```{r foo2, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Selection")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Selection"
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity", "Selection")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Selection"
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Order","Selection")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Selection"
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"

@@

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + geom_line() + theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
```

@@

```{r, foo3, rtmodel-allitems}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lme4::lmer(
  data=dd,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm, method='Wald')[13:20,]
```

@@

```{r, foo4, rtmodeltable, results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1), 
  rtm.summary$coefficients), type='html')
```

@@

```{r, foo5}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0=rtm.confint[,1], y1=rtm.confint[,2], lwd=2)
```

@@

```{r foo6}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r foo7}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber)
```

```{r foo8, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

@@

```{r daz, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

@@

```{r daz2}
dd2 <- subset(dd, abs(rstand)<3.5 )
```

```{r daz3}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lmer(
  data=dd2,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm, method='Wald')
```

@@

```{r daz4,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

@@

```{r daz5}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0=rtm.confint[13:20,1], y1=rtm.confint[13:20,2], lwd=2)
```

@@

```{r daz6}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r daz7}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber
)
```

```{r daz9, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

@@

```{r mc1, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

@@

```{r sep1,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```
